### 2020-07-22 code/figure2.Rmd
fixed scaling and text of figure 2. remove y-axis labels

<div style="text-align: right"> Figure 2 Data </div>
```{r oncoprint-top100-plus-panel-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

@@

```{r figure-2-plot, fig.width=72, fig.height=48, fig.cap="Figure 2: Pan-cancer enrichment of pathway disruptions. Waterfall plot; columns represent tumors. Each row represents a pathway. The pathways are ranked such that the pathway with the most disruptions across all 7,607 tumors is the first row _etc._ and tumors are ordered to prioritize the top pathway disruptions successively. All 377 pathways are shown. Red signifies pathway disrupted pathways, grey signifies not disrupted."}
## top100 pathways pancancer oncoprint
load("../data/onco_ggplot.Rda") ## load data
onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint

fig2A <- ggplot(data = onco_ggplot, aes(x=sample_id, y=path_name)) + 
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  xlab(label = "sample_id") +
  labs(tag = "A") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none")

@@

plots_onco <- list() ## creating a list to store all the ggplots in the for loop below
onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint
for(n in 1:length(clust_col_order)) {
  onco_ggplot <- gdc_reactome_path_analysis %>% .[rownames(.) %in% path_row_order[[n]],] %>% ## only include the top 10 paths
  .[,names(.) %in% clust_col_order[[n]]] %>% ## only include specific tumor class
  rownames_to_column(var = "path_name") %>% ## path_name to col
  gather(., key = "sample_id", value = "mutation_status", -path_name) %>% ## wide to long format
  mutate(path_name = factor(.$path_name, levels = c(rev(path_row_order[[n]]))), ## factor to force oncoprint order
         sample_id = factor(.$sample_id, levels = c(clust_col_order[[n]])))
  plots_onco[[n]] <- ggplot(data = onco_ggplot, aes(x=sample_id, y=path_name)) + ## ggplot code
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  xlab(label = paste0("Class ", n)) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none") 
}

@@

fig2A + plots_onco[[1]] + plots_onco[[2]] + plots_onco[[3]] + plots_onco[[4]] + plots_onco[[5]] + plots_onco[[6]] + plots_onco[[7]] + plots_onco[[8]] + plots_onco[[9]] + plots_onco[[10]] + plot_layout(design = layout) + plot_annotation(tag_levels = "A")
```

>>>>

<div style="text-align: right"> Figure 2 Data </div>
```{r}
library(ggplot2)
library(patchwork)
library(tidyverse)
```

```{r oncoprint-top100-plus-panel-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

@@

```{r figure-2-plot, fig.width=72, fig.height=60, fig.cap="Figure 2: Pan-cancer enrichment of pathway disruptions. Waterfall plot; columns represent tumors. Each row represents a pathway. The pathways are ranked such that the pathway with the most disruptions across all 7,607 tumors is the first row _etc._ and tumors are ordered to prioritize the top pathway disruptions successively. All 377 pathways are shown. Red signifies pathway disrupted pathways, grey signifies not disrupted."}
## top100 pathways pancancer oncoprint
load("../data/onco_ggplot.Rda") ## load data
onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint

fig2A <- ggplot(data = onco_ggplot, aes(x=sample_id, y=path_name)) + 
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 60),
        plot.tag = element_text(size = 48, hjust = 0, vjust = 0),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none") +
  ggtitle("Pan-Cancer")

@@

plots_onco <- list() ## creating a list to store all the ggplots in the for loop below
onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint
for(n in 1:length(clust_col_order)) {
  onco_ggplot <- gdc_reactome_path_analysis %>% .[rownames(.) %in% path_row_order[[n]],] %>% ## only include the top 10 paths
  .[,names(.) %in% clust_col_order[[n]]] %>% ## only include specific tumor class
  rownames_to_column(var = "path_name") %>% ## path_name to col
  gather(., key = "sample_id", value = "mutation_status", -path_name) %>% ## wide to long format
  mutate(path_name = factor(.$path_name, levels = c(rev(path_row_order[[n]]))), ## factor to force oncoprint order
         sample_id = factor(.$sample_id, levels = c(clust_col_order[[n]])))
  plots_onco[[n]] <- ggplot(data = onco_ggplot, aes(x=sample_id, y=path_name)) + ## ggplot code
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 60),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        plot.tag = element_text(size = 48, hjust = 0, vjust = 0),
        legend.position = "none") +
  ggtitle(paste0("Class ", n))
}

@@

fig2A + plots_onco[[1]] + plots_onco[[2]] + plots_onco[[3]] + plots_onco[[4]] + plots_onco[[5]] + plots_onco[[6]] + plots_onco[[7]] + plots_onco[[8]] + plots_onco[[9]] + plots_onco[[10]] + plot_layout(design = layout) + plot_annotation(tag_levels = "A")
```
## oncoprint style of union set pathways
```{r, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

## getting the top 5 pathways from each cluster
n_toppaths <- 5

toppaths <- list() ## list to store data
for(n in 1:max(tumor_clusters$clust_knn)) {
  spec_class <- tumor_clusters %>% filter(clust_knn == n) %>% .$sample_id
  toppaths[[n]] <- gdc_reactome_path_analysis %>% .[, names(.) %in% spec_class] %>% 
    rowSums() %>% as.data.frame() %>%
    'colnames<-' ("Freq") %>%
    rownames_to_column(var = "pathname") %>%
    arrange(desc(Freq)) %>%
    mutate(clust = n) %>% .[1:n_toppaths,]
}

toppaths <- as.data.frame(do.call(rbind, toppaths))
## union set of the top pathways from each cluster
union_toppaths <- unique(toppaths$pathname)

clust_col_order1 <- list()
path_row_order1 <- list()

for(n in 1:max(tumor_clusters$clust_knn)) {
  class_spec_sample_id <- tumor_clusters %>% .[,c("sample_id", "clust_knn")] %>% filter(clust_knn == n) %>% .[,.$sample_id] ## get class spec sample ids
  top10path <- gdc_reactome_path_analysis[union_toppaths, class_spec_sample_id]
  top10path[top10path == "1"] <- "MUT" ## categorize boolean values
  top10path[top10path == "0"] <- " "
  col <- c("MUT" = "firebrick3")
heatmap_legend <- list(title = "", at = c("MUT"),
                       labels = c("Mutation"))
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = "#CCCCCC", col = NA))
    },
    # bug red
    MUT = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), 
            gp = gpar(fill = col["MUT"], col = NA))
    })
  column_title <- "test"
  onco_res <- oncoPrint(top10path,
    alter_fun = alter_fun, col = col,
    column_title = column_title, heatmap_legend_param = heatmap_legend,
    remove_empty_columns = F, remove_empty_rows = F, right_annotation = NULL,
    top_annotation = NULL
    )
  clust_col_order1[[n]] <- names(top10path[,onco_res@column_order])
  path_row_order1[[n]] <- rownames(top10path[onco_res@row_order,])
}

save(clust_col_order1, path_row_order1, file = "../data/onco_clust1.Rda")

## getting tumor sample orders 
for(n in 1) {
df <- as.data.frame(clust_col_order1[[n]])
names(df) <- "sample_id"
df$cluster <- n
for(p in 2:length(clust_col_order1)) {
  df1 <- as.data.frame(clust_col_order1[[p]])
  names(df1) <- "sample_id"
  df1$cluster <- p
  df <- rbind(df, df1)
  rm(df1)
}
}

## setting up for ggplot geom_tile
union_pathways_analysis <- gdc_reactome_path_analysis %>% .[rownames(.) %in% path_row_order1[[1]],] %>% 
  t() %>% as.data.frame() %>% 
  rownames_to_column(var = "sample_id") %>% 
  gather(., key = path_name, value = "mutation_status", -sample_id) %>%
  left_join(., tumor_clusters[,c("sample_id", "clust_knn")]) %>%
  mutate(path_name = factor(.$path_name, levels = c(path_row_order1[[1]])), ## factor to force oncoprint order
         sample_id = factor(.$sample_id, levels = c(rev(df$sample_id))))

onco_color <- c("0" = "grey", "1" = "firebrick3") ## color scheme for oncoprint

fig2_test <- ggplot(data = union_pathways_analysis, aes(y=sample_id, x=path_name)) + ## ggplot code
  geom_tile(aes(fill = as.character(mutation_status))) +
  scale_fill_manual(values = onco_color) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        legend.position = "none") +
  facet_grid(clust_knn ~ ., scales = "free", space = "free", switch = "y")
ggsave("../data/fig2_test.png")

```
